---
title: "Tomato leaves extracts alkaloid profile"
author: "Daniel Quiroz / Cooperstone Lab"
output:
  github_document:
    toc: true
    toc_depth: 4
    number_sections: true
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    number_sections: true
    toc_float: true
    theme: flatly
    code_download: true
---

# What we did

## 1:1 MeOH:H2O water dilution  
Maryam sent samples in 70% methanol, so Daniel diluted them with water 
to be 50% methanol (and 50% water) as that is what we typically analyze
samples in.  He diluted, spun at 21,130 x g for 2 min to pellet any 
insoluble material, and transferred to an LC vial.

## Prepare standars

We also prepared standards for alpha-tomatine 100 ul in 2ml of 
the stock solution (which has dehydrotomatine 
as a contaminant), and tomatidine (which has dehydrotomatidine
as a contaminant). In this situation the contamination is good 
because we get to confirm the presence (or absence) of two more alklaoids.

## Data acquisition

Samples were run on an Agilent 1290 LC interfaced with a 6546 QTOF-MS 
using electrospray ionization and positive ionization mode. 
Separation and MS conditions were the same as in Dzakovich et al.


# Results




* Tutorial about chromatogram plotting 
[here](https://lgatto.github.io/MSnbase/reference/chromatogram-MSnExp-method.html)

```bash
module load R/4.2.0-gnu9.1
```

## Libraries

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Required libraries
library(MSnbase)
library(tidyverse)
library(Rdisop)
library(usethis)
library(readxl)
library(purrr)
library(plotly)

# Utils with core functions
source("utils.R")
```


```{r import_data, cache=TRUE, eval=TRUE, include=FALSE}
mzXML_files <- list.files(pattern = "*.mzXML", path = "./mzXML/", 
                          full.names = T)
  
ms1_data <- readMSData(files = mzXML_files, msLevel. = 1, verbose = F,
                       centroided. = T, mode = "onDisk")
```


```{r eval=TRUE, include=FALSE}
alkaloids <- read_xlsx('./Data/Alkaloids.xlsx', sheet = 1)
alkaloids
```

```{r eval=TRUE, include=FALSE}
alkaloids_short <- alkaloids %>% group_by(Formula) %>% 
  summarize(Alkaloids = str_c(Compound, collapse = "_"))
```


Now, we can calculate the monoisotopic mass and the ranges with a given ppm.

```{r eval=TRUE, include=FALSE}
# Tm = Theoretical mass
# ppm = error 
# range = range of +- ppm
get_ppm <- function(Tm, ppm = 10) {
  Mm = ( ppm * Tm / 1E6) + Tm
  range = abs(Tm-Mm)
  range
}
```


This is the list of the alkaloids with the masses in POS of the analyzed analyties in 
[Dzakovich et al.](https://www.frontiersin.org/articles/10.3389/fpls.2020.00767/full?&utm_source=Email_to_authors_&utm_medium=Email&utm_content=T1_11.5e1_author&utm_campaign=Email_publication&field=&journalName=Frontiers_in_Plant_Science&id=523863)

```{r eval=TRUE}
alkaloids_short$Monoisotopic <- sapply(alkaloids_short$Formula, 
                                       function(x) getMolecule(x)$exactmass)
alkaloids_short <- alkaloids_short %>% 
  mutate(mz_pos = Monoisotopic + 1.007) %>% 
  mutate(Upper_limit = mz_pos + get_ppm(mz_pos, ppm = 10),
         Lower_limit = mz_pos - get_ppm(mz_pos, ppm = 10) )
alkaloids_short
```


## Samples are concentrated enough

The samples are plenty concentrated enough, we injected 0.1 ul (the minimum we can)
and were nearly saturating signal on our MS. This is good because we can always dilute if
necessary (but concentrating is much harder).

## Alpha-tomatine and dehydromatidine in your samples

We can confirm by comparing with authentic standard that there is alpha-tomatine and tomatidine in your samples


### Alpha-tomatine (C50H83NO21) rt = 5.4

```{r cache=TRUE, eval=TRUE, include=TRUE}
aTomatine <- get_EIC_tbb(msn_exp = ms1_data, mz = c(1034.5424, 1034.5631), ID = "C50H83NO21")
```



```{r fig.height=8, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
aTomatine_plot <- ggplot(aTomatine, aes(rt/60, Intensity, col = Sample_ID)) +
  geom_line() + facet_wrap("Sample_ID", ncol = 1, scales = "free_y") + guides(color = "none") + 
  theme_bw() + 
  labs(title = "Alpha-tomatine EIC ([M+H]+ = 1034.5528)",
       subtitle = "C50H83NO21: Alpha-tomatine\nExpected retention time (rt): 5.4",
       x = "rt (min)") +
  geom_vline(xintercept = 5.35, alpha =0.6, linetype = "dashed")
aTomatine_plot
```

**Interactive plots** 

```{r echo=FALSE, fig.height=8, fig.width=8}
line_specs <- list(width = 1.3)

AT_std <- aTomatine %>% dplyr::filter(Sample_ID == "aT_tomatidine_std_0pt1ul_4.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "Standard", type = "scatter", mode = "lines", line = line_specs)

AT_SC22 <- aTomatine %>% dplyr::filter(Sample_ID == "SC22_1_0pt1ul_3.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "SC22", type = "scatter", mode = "lines",  line = line_specs)


AT_SCR02 <- aTomatine %>% dplyr::filter(Sample_ID == "SCR02_1_0pt1ul_2.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "SCR02", type = "scatter", mode = "lines",  line = line_specs)

subplot(AT_std, AT_SC22, AT_SCR02, nrows = 3, shareX = T) %>% 
  layout(title = list(text = "Alpha-Tomatine chromatograms (rt: 5.4)")) 
```



### Dehydrotomatine (Dehydrotomatine) rt = 5.2


```{r cache=TRUE, eval=TRUE, include=TRUE}
DOHtine <- get_EIC_tbb(msn_exp = ms1_data, mz = c(1032.5268, 1032.5474), ID = "C50H81NO21")
```



```{r fig.height=8, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
DOHtine_plot <- ggplot(DOHtine, aes(rt/60, Intensity, col = Sample_ID)) +
  geom_line() + facet_wrap("Sample_ID", ncol = 1, scales = "free_y") + guides(color = "none") + 
  theme_bw() + 
  labs(title = "Dehydrotomatine EIC ([M+H]+ = 1032.5)",
       subtitle = "C50H81NO21: Dehydrotomatine\nExpected retention time (rt): 5.2",
       x = "rt (min)") +
  geom_vline(xintercept = 5.2, alpha = 0.3, linetype = "dashed")
DOHtine_plot
```

**Interactive Plots**


```{r echo=FALSE, fig.height=8, fig.width=8}
DOHtine_std <- DOHtine %>% dplyr::filter(Sample_ID == "aT_tomatidine_std_0pt1ul_4.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "Standard", type = "scatter", mode = "lines", line = line_specs)

DOHtine_SC22 <- DOHtine %>% dplyr::filter(Sample_ID == "SC22_1_0pt1ul_3.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "SC22", type = "scatter", mode = "lines",  line = line_specs)


DOHtine_SCR02 <- DOHtine %>% dplyr::filter(Sample_ID == "SCR02_1_0pt1ul_2.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "SCR02", type = "scatter", mode = "lines",  line = line_specs)

subplot(DOHtine_std, DOHtine_SC22, DOHtine_SCR02, nrows = 3, shareX = T) %>% 
  layout(title = list(text = "Alpha-Tomatine chromatograms (rt: 5.2)")) 
```


## We do not see Tomatidine or Dehydrotomatidine

### Tomatidine (C27H45NO2) rt = 7

```{r cache=TRUE, eval=TRUE, include=TRUE}
tomatidine <- get_EIC_tbb(msn_exp = ms1_data, mz = c(416.3562, 416.3479), ID = "C27H45NO2")
```



```{r fig.height=8, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
tomatidine_plot <- ggplot(tomatidine, aes(rt/60, Intensity, col = Sample_ID)) +
  geom_line() + facet_wrap("Sample_ID", ncol = 1, scales = "free_y") + guides(color = "none") + 
  theme_bw() + 
  labs(title = "Tomatidine EIC ([M+H]+ = 416.4)",
       subtitle = "C27H45NO2: Tomatidine\nExpected retention time (rt): 7.0",
       x = "rt (min)") +
  geom_vline(xintercept = 6.9, alpha = 0.3, linetype = "dashed")
tomatidine_plot
```

**Interactive Plots**

```{r echo=FALSE, fig.height=8, fig.width=8}
TO_std <- tomatidine %>% dplyr::filter(Sample_ID == "aT_tomatidine_std_0pt1ul_4.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "Standard", type = "scatter", mode = "lines",  line = line_specs)

TO_SC22 <- tomatidine %>% dplyr::filter(Sample_ID == "SC22_1_0pt1ul_3.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "SC22", type = "scatter", mode = "lines",  line = line_specs)


TO_SCR02 <- tomatidine %>% dplyr::filter(Sample_ID == "SCR02_1_0pt1ul_2.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "SCR02", type = "scatter", mode = "lines",  line = line_specs)

subplot(TO_std, TO_SC22, TO_SCR02, nrows = 3, shareX = T) %>% 
  layout(title = list(text = "Tomatidine chromatograms (rt = 7)")) 
```




### Dehydrotomatidine (C27H43NO2) rt: 7.3


```{r cache=TRUE, eval=TRUE, include=TRUE}
DOHdine <- get_EIC_tbb(msn_exp = ms1_data, mz = c(414.3322, 414.3405), ID = "C27H43NO2")
```


```{r fig.height=8, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
DOHdine_plot <- ggplot(DOHdine, aes(rt/60, Intensity, col = Sample_ID)) +
  geom_line() + facet_wrap("Sample_ID", ncol = 1, scales = "free_y") + guides(color = "none") + 
  theme_bw() + 
  labs(title = "Dehydrotomatidine EIC ([M+H]+ = 414.3)",
       subtitle = "C27H43NO2: Dehydrotomatidine\nExpected retention time (rt): 7.3",
       x = "rt (min)") +
  geom_vline(xintercept = 7.3, alpha = 0.3, linetype = "dashed")
DOHdine_plot
```

**Interactive Plots**


```{r echo=FALSE, fig.height=8, fig.width=8}
DOHdine_std <- DOHdine %>% dplyr::filter(Sample_ID == "aT_tomatidine_std_0pt1ul_4.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "Standard", type = "scatter", mode = "lines", line = line_specs)

DOHdine_SC22 <- DOHdine %>% dplyr::filter(Sample_ID == "SC22_1_0pt1ul_3.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "SC22", type = "scatter", mode = "lines",  line = line_specs)


DOHdine_SCR02 <- DOHdine %>% dplyr::filter(Sample_ID == "SCR02_1_0pt1ul_2.mzXML") %>% 
  plot_ly(x = ~rt/60, y = ~Intensity, name = "SCR02", type = "scatter", mode = "lines",  line = line_specs)

subplot(DOHdine_std, DOHdine_SC22, DOHdine_SCR02, nrows = 3, shareX = T) %>% 
  layout(title = list(text = "Dehydrotomatidine chromatograms (rt: 7.3)")) 
```

# Usefull links
* msnBase chromatogram plotting [link](https://lgatto.github.io/MSnbase/articles/v03-MSnbase-centroiding.html)


# Session info

```{r}
sessionInfo()
```

